### 2024-02-11 paper/paper.qmd
Add files via upload

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(dplyr)
library(stringr)
library(ggplot2)
library(MetBrewer)

```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

# Results {#sec-result}

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| 
#### Loading the Cleaned Dataset ###

lgbtq <- read_csv("data/analysis_data/lgbtq.csv")
billion <- read_csv("data/analysis_data/billion.csv")
partisans <- read_csv("data/analysis_data/partisans.csv")
```

```{r}
#| echo: false
#| eval: true
#| label: prepare_lgbtq
#| warning: false

#### Calculate the percentages of handles given by people with different gender, race and education level ####
### in LGBTQ target ###
### Prepare for later graphs ###

#gender
counts <- lgbtq %>%
  group_by(treatment,gender,handle) %>%
  summarise(count = n(), .groups = "drop") 
print(counts)

total_counts <- lgbtq %>%
  group_by(treatment, gender) %>%
  summarise(total = n(), .groups = "drop") %>%
  ungroup()

percentages <- counts %>%
  left_join(total_counts, by = c("treatment", "gender")) %>%
  mutate(percentage = (count / total) * 100)

result_l_gender <- percentages %>%
  filter(gender != "Other") %>% #Number of Others is too small, delete it for simplicity
  dplyr::select(treatment, gender, handle, percentage)

#race
counts_race <- lgbtq %>%
  filter(!is.na(racecat_anes)) %>%
  group_by(treatment,racecat_anes,handle) %>%
  summarise(count = n(), .groups = "drop")
print(counts_race)

total_counts_race <- lgbtq %>%
  filter(!is.na(racecat_anes)) %>%
  group_by(treatment, racecat_anes) %>%
  summarise(total = n(), .groups = "drop") %>%
  ungroup()

print(total_counts_race)
percentages_race <- counts_race %>%
  left_join(total_counts_race, by = c("treatment", "racecat_anes")) %>%
  mutate(percentage = (count / total) * 100)

result_l_race <- percentages_race %>%
  dplyr::select(treatment, racecat_anes, handle, percentage)

#education
counts_edu <- lgbtq %>%
  filter(!is.na(educ)) %>%
  group_by(treatment,educ,handle) %>%
  summarise(count = n(), .groups = "drop")

total_counts_edu <- lgbtq %>%
  filter(!is.na(educ)) %>%
  group_by(treatment, educ) %>%
  summarise(total = n(), .groups = "drop") %>%
  ungroup()

percentages_edu <- counts_edu %>%
  left_join(total_counts_edu, by = c("treatment", "educ")) %>%
  mutate(percentage = (count / total) * 100)

result_l_edu <- percentages_edu %>%
  filter(educ %in% c("High school graduate", "College","Postgraduate (e.g. Masters)"))%>%
  dplyr::select(treatment, educ, handle, percentage) 
```

```{r}
#| echo: false
#| eval: true
#| label: prepare_billionaire
#| warning: false

#gender
count <- billionaires %>%
  group_by(treatment,gender,handle) %>%
  summarise(count = n(), .groups = "drop")
print(count)

total_counts <- billionaires %>%
  group_by(treatment, gender)%>%
  summarise(total = n(), .groups = "drop") %>%
  ungroup()

percentages <- count %>%
  left_join(total_counts, by = c("treatment", "gender")) %>%
  mutate(percentage = (count / total) * 100)

result_b_gender <- percentages %>%
  filter(gender != "Other") %>%
  dplyr::select(treatment, gender, handle, percentage)

print(result_b_gender)

#race
counts_race <- billionaires %>%
  filter(!is.na(racecat_anes)) %>%
  group_by(treatment,racecat_anes,handle) %>%
  summarise(count = n(), .groups = "drop")
print(counts_race)

total_counts_race <- billionaires %>%
  filter(!is.na(racecat_anes)) %>%
  group_by(treatment, racecat_anes) %>%
  summarise(total = n(), .groups = "drop") %>%
  ungroup()

percentages_race <- counts_race %>%
  left_join(total_counts_race, by = c("treatment", "racecat_anes")) %>%
  mutate(percentage = (count / total) * 100)

result_b_race <- percentages_race %>%
  dplyr::select(treatment, racecat_anes, handle, percentage)

#education
counts_edu <- billionaires %>%
  filter(!is.na(educ)) %>%
  group_by(treatment,educ,handle) %>%
  summarise(count = n(), .groups = "drop")

total_counts_edu <- billionaires %>%
  filter(!is.na(educ)) %>%
  group_by(treatment, educ) %>%
  summarise(total = n(), .groups = "drop") %>%
  ungroup()

print(total_counts_edu)
percentages_edu <- counts_edu %>%
  left_join(total_counts_edu, by = c("treatment", "educ")) %>%
  mutate(percentage = (count / total) * 100)

result_b_edu <- percentages_edu %>%
  filter(educ %in% c("High school graduate", "College","Postgraduate (e.g. Masters)"))%>%
  dplyr::select(treatment, educ, handle, percentage)

```

```{r}
#| echo: false
#| eval: true
#| label: gender comparision
#| warning: false

create_gender_comparison_plot <- function(data, title) {
  levels_treatment <- unique(data$treatment)
  levels_gender <- c("Female", "Male")

  # Create a new factor combining treatment and gender with the right order
  data$treatment_gender <- factor(
    paste(data$treatment, data$gender),
    levels = c(rbind(
      paste(levels_treatment, "Female"),
      paste(levels_treatment, "Male")
    ))
  )

  # We'll create a named vector to map the factor levels to the custom labels
  # Initialize the labels with treatment names, ensuring they only appear once
  treatment_labels <- rep("", length(levels_treatment) * 2)
  names(treatment_labels) <- levels(data$treatment_gender)
  treatment_labels[names(treatment_labels) %in% paste(levels_treatment, "Female")] <- paste("\n", levels_treatment)

  # Now, assign the gender labels, making sure they're in the first line
  treatment_labels[names(treatment_labels) %in% paste(levels_treatment, "Female")] <- paste("Female", treatment_labels[names(treatment_labels) %in% paste(levels_treatment, "Female")])
  treatment_labels[names(treatment_labels) %in% paste(levels_treatment, "Male")] <- "Male"

  ggplot(data, aes(x = treatment_gender, y = percentage, fill = handle)) +
    geom_bar(stat = "identity", position = position_fill()) +
    scale_y_continuous(labels = scales::percent_format()) +
    labs(x = "", y = "Percentage", fill = "How should social media companies \nhandle the post?", title = title) + # Updated legend label
    scale_x_discrete(labels = treatment_labels) + # Use custom labels for the x axis
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 0, hjust = 0.5, vjust = 1), # Adjust text alignment if needed
          legend.position = "bottom",
          legend.box = "vertical",
          legend.key.size = unit(0.5, 'cm')) +
    guides(fill = guide_legend(reverse = TRUE, nrow = 3, byrow = TRUE)) + # Adjust the number of rows and order
    scale_fill_met_d(name = "Degas", direction = -1) # Apply the Degas palette from MetBrewer
}

# Then you can create and display the plots as before
lgbtq_plot <- create_gender_comparison_plot(result_l_gender, "LGBTQ Topic: Gender Comparison by Treatment")
billionaire_plot <- create_gender_comparison_plot(result_b_gender, "Billionaire Topic: Gender Comparison by Treatment")
print(lgbtq_plot)
print(billionaire_plot)

```

```{r}
#| echo: false
#| eval: true
#| label: race comparision
#| warning: false

create_race_comparison_plot <- function(data, title) {
  levels_treatment <- unique(data$treatment)
  levels_race <- c("Black", "Hispanic", "White", "Race other/multiple")

  # Create a new factor combining treatment and race with the right order
  data$treatment_race <- factor(
    paste(data$treatment, data$racecat_anes),
    levels = c(rbind(
      paste(levels_treatment, "Black"),
      paste(levels_treatment, "Hispanic"),
      paste(levels_treatment, "White"),
      paste(levels_treatment, "Race other/multiple")
    ))
  )

  # We'll create a named vector to map the factor levels to the custom labels
  # Initialize the labels with treatment names, ensuring they only appear once
  treatment_labels <- rep("", length(levels_treatment) * length(levels_race))
  names(treatment_labels) <- levels(data$treatment_race)
  treatment_labels[names(treatment_labels) %in% paste(levels_treatment, "Black")] <- paste("\n", levels_treatment)

  # Assign the race labels, making sure they're in the first line
  treatment_labels[names(treatment_labels) %in% paste(levels_treatment, "Black")] <- paste("Black", treatment_labels[names(treatment_labels) %in% paste(levels_treatment, "Black")])
  treatment_labels[names(treatment_labels) %in% paste(levels_treatment, "Hispanic")] <- "Hispanic"
  treatment_labels[names(treatment_labels) %in% paste(levels_treatment, "White")] <- "White"
  treatment_labels[names(treatment_labels) %in% paste(levels_treatment, "Race other/multiple")] <- "Race other/multiple"

  ggplot(data, aes(x = treatment_race, y = percentage, fill = handle)) +
    geom_bar(stat = "identity", position = position_fill()) +
    scale_y_continuous(labels = scales::percent_format()) +
    labs(x = "", y = "Percentage", fill = "How should social media companies \nhandle the post?", title = title)+
    scale_x_discrete(labels = treatment_labels) + 
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "bottom",
          legend.box = "vertical",
          legend.key.size = unit(0.5, 'cm')) +
    guides(fill = guide_legend(reverse = TRUE, nrow = 3, byrow = TRUE)) +
    scale_fill_met_d(name = "Degas", direction = -1) 
}

lgbtq_plot <- create_race_comparison_plot(result_l_race, "LGBTQ Topic: Race Comparison by Treatment")
billionaire_plot <- create_race_comparison_plot(result_b_race, "Billionaire Topic: Race Comparison by Treatment")

print(lgbtq_plot)
print(billionaire_plot)

```

```{r}
#| echo: false
#| eval: true
#| label: education level comparision
#| warning: false

create_edu_comparison_plot <- function(data, title) {
  levels_treatment <- unique(data$treatment)
  levels_edu <- c("High school graduate", "College", "Postgraduate (e.g. Masters)")

  # Create a new factor combining treatment and education with the right order
  data$treatment_edu <- factor(
    paste(data$treatment, data$educ),
    levels = c(rbind(
      paste(levels_treatment, "High school graduate"),
      paste(levels_treatment, "College"),
      paste(levels_treatment, "Postgraduate (e.g. Masters)")
    ))
  )

  # We'll create a named vector to map the factor levels to the custom labels
  # Initialize the labels with treatment names, ensuring they only appear once
  treatment_labels <- rep("", length(levels_treatment) * length(levels_edu))
  names(treatment_labels) <- levels(data$treatment_edu)
  treatment_labels[names(treatment_labels) %in% paste(levels_treatment, "High school graduate")] <- paste("\n", levels_treatment)

  # Now, assign the education labels, making sure they're in the first line
  treatment_labels[names(treatment_labels) %in% paste(levels_treatment, "High school graduate")] <- paste("High school graduate", treatment_labels[names(treatment_labels) %in% paste(levels_treatment, "High school graduate")])
  treatment_labels[names(treatment_labels) %in% paste(levels_treatment, "College")] <- "College"
  treatment_labels[names(treatment_labels) %in% paste(levels_treatment, "Postgraduate (e.g. Masters)")] <- "Postgraduate (e.g. Masters)"

  ggplot(data, aes(x = treatment_edu, y = percentage, fill = handle)) +
    geom_bar(stat = "identity", position = position_fill()) +
    scale_y_continuous(labels = scales::percent_format()) +
    labs(x = "", y = "Percentage", fill = "How should social media companies \nhandle the post?", title = title) + # Updated legend label
    scale_x_discrete(labels = treatment_labels) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "bottom",
          legend.box = "vertical",
          legend.key.size = unit(0.5, 'cm')) +
    guides(fill = guide_legend(reverse = TRUE, nrow = 3, byrow = TRUE)) + 
    scale_fill_met_d(name = "Degas", direction = -1) 
}

# Create and display the plots separately
lgbtq_plot <- create_edu_comparison_plot(result_l_edu, "LGBTQ Topic: Education Comparison by Treatment")
billionaire_plot <- create_edu_comparison_plot(result_b_edu, "Billionaire Topic: Education Comparison by Treatment")

print(lgbtq_plot)
print(billionaire_plot)

```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

\newpage

# References